---
title: "J20_EMG_Analysis"
author: "Iris Oren"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: default
  pdf_document: default
---
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Description
This .rmd file analyses simultaneously recorded EEG and EMG data to assess whether EMG power is reduced during period of high theta/delta during sleep (ie. are high theta periods indicative of REM or ectopic theta)

*Data
Dual EEG and EMG data from 4 animals analysed using SCPP5 processor. 

SCPP5.tcl was run on 8s epochs of Data, with glitch=200. The processor generated text files, : "ndfFilename_SCPP5V1.txt", organised in individual folders by AnimalID

(Raw Data: Dropbox/Analysis/Neuroarchiver/J20EEG/SCPP5V1).

Processor output for 8s epoch
Filename IntervalTime Loss Spikes SpikeBursts DeltaPower ThetaPower EMGPower

###Package Initialisation 

First, load packages and source files
```{r LoadPackages_Chunk, echo=FALSE, include=FALSE}
library(lubridate)
library(plyr)
library(Rmisc)
library(lazyeval) #Needed for using ggplot2 library
library(ggplot2)
library(grid)
library(cowplot)
library(stringr)
library(scatterplot3d)
library(data.table)
library(perm)
library(circular)
library(plotrix)
library(readr)  #for col_skip in read.csv
library(lme4)
library(lsmeans)
library(multcomp)
library(CircOutlier)
library(dplyr)


sourceDir <- function(path, trace = TRUE, ...) {
    for (nm in list.files(path, pattern = "\\.[Rr]$")) {
       if(trace) cat(nm,":")           
       source(file.path(path, nm), ...)
       if(trace) cat("\n")
    }
}
sourceDir("R/")  #All function in R/ folder relative to project
```
### Datafile and variable initialisation
Specify files and folders to be used
```{r VariableInitialisationChunk, ECHO=FALSE}
NDFDataParentFolder<-"~/Dropbox/ANALYSIS/Neuroarchiver/J20EEG/SCPP5V1"
RecordingInfoFile<-"./Data/RecordingInfoEMG.csv"
FalsePositiveFile<-"./Data/SCPP5V1_Performance.csv"
LossIntervalPercentageFile<-"./Data/LossIntervalPercentageEMG.csv"

# RecordingInfoFile<-"./Data/RecordingInfo_WT_VideoScoring.csv"
OutputFolder<-"./Output"
```


```{r ImportRecordingInformation, ECHO=FALSE, include=TRUE}
detach("package:dplyr", character.only = TRUE)
library("dplyr", character.only = TRUE)

FalsePositiveCutOff=0.1
LossIntervalPercentageThreshold<-5 #Limit for percent of intervals lost to include animal in analysis

SummaryDataframe<-read.csv(RecordingInfoFile, na = "empty")
SummaryDataframe$AnimalID<-as.character(SummaryDataframe$AnimalID)
SummaryDataframe$TimeZone<-as.character(SummaryDataframe$TimeZone)

```


```{r ImportAndTransformNDFOutput_chunk, ECHO=FALSE}
if(file.exists("./Output/SCPPDataframeEMGOriginal.csv")){
  SCPPDataframeEMG<-read.csv("./Output/SCPPDataframeEMG.csv")
  RemoveOption<-0
  } else{  #Else create SCPPDataframeEMG.csv
    
      #Initialise variables all animals
     # SCPPDataframeList<-list()
      SCPPDataframeEMG<-data.frame()
      
      for(i in 1:nrow(SummaryDataframe)){     #Loop through animals
      
        SCPPDataframe<-data.frame() #Initialise dataframe for current animal
        
       #Write AnimalID, genotype and TimeZone to variable for current animal
              AnimalIDCurrent<-SummaryDataframe$AnimalID[i]
              GenotypeCurrent<-SummaryDataframe$Genotype[i]
              TimeZoneCurrent<-SummaryDataframe$TimeZone[i]
              
        #i. Set AnimalID subfolder
        
          FullPath<-paste(NDFDataParentFolder, AnimalIDCurrent, sep="/")
       FullPath<-paste(FullPath, "/", sep="")
       #Next import all files in input_path to create a new r dataframe called  `SCPPDataframe'
          SCPPDataframe<-ImportData(FullPath)
        
          

           # Exclude Loss >20 % and Artifact intervals which have delta power >1000ksq counts and write the number of ExcludedIntervals to ExcludedIntervalsVariable in SCPPDataframe
          LossThreshold=20
          DeltaThreshold=1000  
          
         
# Add Variable for ExcludeInterval: If Loss>LossThreshold or  delta power>DeltaThreshold
          
ExcludedIntervalsVariable<-which((SCPPDataframe$Loss>LossThreshold)| (SCPPDataframe$Delta>DeltaThreshold))
#Set ExcludeInterval to a default of 0
          SCPPDataframe$ExcludeInterval<-0
#ExcludeInterval<-1 for exclusion
          SCPPDataframe$ExcludeInterval[ExcludedIntervalsVariable]<-1
            

          
#Apend time information to Data frame  
          SCPPDataframe <- GetTimeOfDayAndDate(SCPPDataframe, TimeZoneCurrent)
# SCPPDataframe$TimeOfDay<-as.POSIXct(SCPPDataframe$TimeOfDay, tz=TimeZoneCurrent)
          
           
#Apend AnimalID variable and Genotype
          AnimalIDVariable<-rep(AnimalIDCurrent, nrow(SCPPDataframe))
          SCPPDataframe<-cbind(SCPPDataframe, AnimalIDVariable)
          Genotype<-rep(GenotypeCurrent, nrow(SCPPDataframe))
          SCPPDataframe<-cbind(SCPPDataframe, Genotype)
      
#Apend SCPPDataframe to SCPPDataframeEMG. To access an animal, used "filter()" 
          SCPPDataframeEMG<-rbind(SCPPDataframeEMG, SCPPDataframe)
}
  
# Add LightDark variable where LightDark=TRUE is lights on
    Hour<-lubridate::hour(SCPPDataframeEMG$TimeOfDay)
    SCPPDataframeEMG$LightDark <- Hour %in% seq(7, 18)
    
# Add ThetaOverDelta
    SCPPDataframeEMG$ThetaOverDelta<-SCPPDataframeEMG$Theta/SCPPDataframeEMG$Delta
       
# Rename SpikeCount to Single spikecount and calculate total spike count as burst + single
    SCPPDataframeEMG$SingleSpikeCount<-SCPPDataframeEMG$SpikeCount
    SCPPDataframeEMG$SpikeCount<-SCPPDataframeEMG$SingleSpikeCount+SCPPDataframeEMG$BurstCount
    

  RemoveOption<-1
    }
# Exclude intervals
SCPPDataframeEMG<-filter(SCPPDataframeEMG, ExcludeInterval==0)

if(RemoveOption==1){
    rm(SCPPDataframe, FalsePositives, LossIntervalPercentage, Treatment, Genotype)
}
```

Import VideoScoring dataframe and amend with UnixTime 
```{r MakeVideoScoringDataframe, echo=FALSE, include=FALSE}

FileList<-c("./Data/VideoScoring/VideoScoringJ0527.csv",
            "./Data/VideoScoring/VideoScoringJ0528.csv",   
            "./Data/VideoScoring/VideoScoringJ0530.csv",  
            "./Data/VideoScoring/VideoScoringJ0533.csv"
            )


AnimalIDList<-c("J0527", "J0528", "J0530", "J0533")
SCPPDataframeVideoTimesEMG<-data.frame()
AnimalIDVideoStartEnd<-list() #A list to store Start and End times for video data. The list index is AnimalID_FileIndex eg. JF220_1

for(FileIndex in 1:length(FileList)){
 
   VideoScoringFilename<-FileList[[FileIndex]] 
  
   AnimalIDCurrent<-AnimalIDList[[FileIndex]]
   
    #Import
    VideoScoring<- read.csv(VideoScoringFilename)
    VideoScoring<-dplyr::select(VideoScoring, DateOfEEG, TimeOfEEG,  SleepWake)
    VideoScoring$SleepWake<-as.character(VideoScoring$SleepWake) #needs to be character. "as.factor" converts to numeric factors
    
    #Add date to Time to make TimeString
    VideoScoring$DateTime<-paste(VideoScoring$DateOfEEG, VideoScoring$TimeOfEEG)
    
    VideoScoring$DateTime<-strptime(as.character(VideoScoring$DateTime), "%d/%m/%Y %H:%M:%S")
    
    #Convert to UnixTime
    VideoScoring$UnixTime<-(as.numeric(as.POSIXct(VideoScoring$DateTime)))
    #VideoScoring<-VideoScoring[complete.cases(VideoScoring)]
    #Determine Start and End time of video file for particular day
    VideoStartTime<-VideoScoring$UnixTime[1]
    VideoEndTime<-VideoScoring$UnixTime[nrow(VideoScoring)]
    
    #Create vector with videostart and videoend variables for this FileIndex iteration
    AnimalIDVideoStartEndVectorCurrent<-c(as.numeric(VideoStartTime), as.numeric(VideoEndTime))
    
    ListIndex<-paste(AnimalIDCurrent, as.character(FileIndex), sep="_")
    AnimalIDVideoStartEnd[[ListIndex]]<-AnimalIDVideoStartEndVectorCurrent
    
    #Filter SCPP1 between start and endtime of video for particular day

  SCPPDataframeVideoTimes<-filter(SCPPDataframeAll,
                                  AnimalIDVariable==AnimalIDCurrent)
  SCPPDataframeVideoTimes<-filter(SCPPDataframeVideoTimes,
                                  InitialisedTime>=VideoStartTime)
    SCPPDataframeVideoTimes<-filter(SCPPDataframeVideoTimes, InitialisedTime<=VideoEndTime)
    
    
    #Add video score to dataframe for which video is available
    
    SCPPDataframeVideoTimes$SleepWake<-AddVideoScoreToSCPPData(SCPPDataframeVideoTimes$InitialisedTime, VideoScoring$UnixTime, VideoScoring$SleepWake)
    SCPPDataframeVideoTimesAll<-rbind(SCPPDataframeVideoTimesAll, SCPPDataframeVideoTimes)
  
    SCPPDataframeVideoTimesAll$SleepWake<-as.factor(SCPPDataframeVideoTimesAll$SleepWake)
}

rm(SCPPDataframeVideoTimes, VideoScoring)
```


Filter for sleep episodes identified by video and add REM/NREM variable
```{r REM_NREM, echo=FALSE}

SCPPDataframeSleep<-filter(SCPPDataframeVideoTimesAll, SleepWake=="S")
SCPPDataframeSleep$ThetaDeltaHiLo<-ifelse(ThetaOverDelta>2, "High", 
                                                  ifelse(ThetaOverDelta<1, "Low", "U"))
                                                  
SCPPDataframeSleep<-filter(SCPPDataframeSleep, ThetaDeltaHiLo=="High" | ThetaDeltaHiLo=="Low")
```

Compare EMG power in periods with theta/delta high and low
```{r, EMG_Compare_PlottingChunk, include=TRUE, echo=FALSE}
#if("plyr" %in% (.packages())){
 # detach("package:plyr", unload=TRUE)
#}
#Group by AnimalID and ThetaDeltaHiLo
ThetaDeltaHiLo_grouped<-dplyr::group_by(SCPPDataframeSleep, AnimalIDVariable, ThetaDeltaHiLo)
#Compute the mean EMG Power for each animal in each behavioural state                     
SummaryEMG<-dplyr::summarise(ThetaDeltaHiLo_grouped,
                     MeanEMG=mean(EMG),
                     num=n(),
                     SEM=sd(EMG)/sqrt(n()),
                     CI_L=CI(EMG)[1],
                     CI_U=CI(EMG)[3]
                     )




EMG_Compare_Plot<-ggplot(SummaryEMG, 
               aes(x=REM_NREM, 
                   y=MeanEMG)) +
  geom_line(aes(group=AnimalIDVariable)) +
  geom_point(aes(group=AnimalIDVariable)) +
  geom_errorbar(aes(ymin=CI_L, ymax=CI_U, width=0.4))+
  ylab("EMG power") +
  xlab(expression(theta) "/" expression(delta)) +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle=60, hjust=1))#+
  #scale_color_brewer(palette = "Set1")
  


EMGPlot

```